---
title: "Georgia COVID-19 Vaccination Rates"
author: Amaka Mgboh
date: 03/25/2025
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    theme: cerulean
  html_document:
    df_print: paged
    self_contained: true
---

```{r setup, echo=FALSE, eval = TRUE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message= FALSE)
##options(repos = c(CRAN = "https://cran.rstudio.com"))
##install.packages("readxl")
##Sys.setenv(RSTUDIO_PANDOC = "/opt/homebrew/bin/pandoc")
```

## Introduction
This report explores geographic disparities in COVID-19 vaccination rates across Georgia counties. It examines the influence of socioeconomic factors and healthcare access on vaccine distribution using interactive visualizations.

```{r libraries, message=FALSE, warning=FALSE, echo=FALSE, eval = TRUE}
library(ggplot2)
library(plotly)
library(leaflet)
library(dplyr)
library(sf)
```

## Load Data
```{r load-data, eval = TRUE}
file_path <- "Home/Downloads/Georgia_DPH_PUBLIC_Vaccination_Public_Data_in_Excel.xlsx"
# install.packages("readxl")  
library(readxl)
df <- read_excel("~/Downloads/Georgia_DPH_PUBLIC_Vaccination_Public_Data_in_Excel.xlsx", 
                 sheet = "COUNTY_SUMMARY")
```

```{r}
# Load libraries
library(sf)
library(tigris)
library(dplyr)
library(plotly)
library(DT)
library(readr)
library(leaflet)
library(RColorBrewer)
library(flexdashboard)

# Step 1: Get Georgia county shapefiles
options(tigris_use_cache = TRUE)
ga_sf <- counties(state = "GA", cb = TRUE, class = "sf")


colnames(ga_sf)  
colnames(df)     

# Step 1: Clean and prepare ga_data
ga_data <- df %>%
  select(COUNTY_ID, PCTCUMPVAX)  

# Step 2: Perform the left join to merge ga_data into ga_sf
ga_sf <- merge(ga_sf, ga_data, by.x="GEOID", by.y="COUNTY_ID")
ga_sf$PCTCUMPVAX = as.numeric(ga_sf$PCTCUMPVAX)
ga_sf$PCTCUMPVAX = ga_sf$PCTCUMPVAX / 100 

# Step 4: Define color palette
pal <- colorNumeric(palette = "YlGnBu", domain = ga_sf$PCTCUMPVAX)

# Step 5: Plot with leaflet
leaflet(ga_sf) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(
    fillColor = ~pal(PCTCUMPVAX),
    weight = 1,
    color = "white",
    fillOpacity = 0.7,
    label = ~paste0(NAME, ": ", round(PCTCUMPVAX, 2)),
    highlightOptions = highlightOptions(weight = 2, color = "#666", fillOpacity = 0.9, bringToFront = TRUE)
  ) %>%
  addLegend(pal = pal, values = ~PCTCUMPVAX, title = "Vaccination Rate %", position = "bottomright")


```


### Key Takeaways:
- Identifies counties with the lowest vaccination rates.
- Shows correlation between vaccination rates and healthcare facility distribution.
- Highlights potential target areas for intervention.

## Table: Vaccination Rate by District Name
```{r scatterplot, message=FALSE, warning=FALSE}
library(ggplot2)
library(plotly)
library(scales)
library(DT) 

df$PCTCUMPCVAX <- as.numeric(df$PCTCUMPCVAX)

# Create scatter plot with facet_wrap
plot <- ggplot(df, aes(x = COUNTY_NAME, y = PCTCUMPCVAX / 100, text = COUNTY_NAME)) + 
  geom_point(color = "blue", alpha = 0.7) +
  labs(title = "Vaccination Rate by District Name",
       x = "District Name",
       y = "Vaccination Rate (%)") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8), 
        axis.ticks.y = element_blank()) + 
  scale_y_continuous(labels = percent_format(scale = 1),  
                     breaks = seq(0, 1, by = 0.1)) + 
  facet_wrap(~DISTRICT_NAME) 

# Make the plot interactive
plotly_plot <- ggplotly(plot, tooltip = c("text", "x", "y"))

# Create a data table
datatable(df, options = list(pageLength = 10, autoWidth = TRUE))

```

```{r}
# Load libraries
library(sf)
library(tigris)
library(dplyr)
library(leaflet)
library(RColorBrewer)

# Step 1: Get Georgia county shapefiles
options(tigris_use_cache = TRUE)
ga_sf <- counties(state = "GA", cb = TRUE, class = "sf")

# Step 2: Clean and prepare ga_data
ga_data <- df %>%
  select( COUNTY_ID, PCTCUMPVAX)  

# Step 3: Perform the left join to merge ga_data into ga_sf
ga_sf <- merge(ga_sf, ga_data, by.x = "GEOID", by.y = "COUNTY_ID")
ga_sf$PCTCUMPVAX <- as.numeric(ga_sf$PCTCUMPVAX)
ga_sf$PCTCUMPVAX <- ga_sf$PCTCUMPVAX / 100  

# Step 4: Define the color palette
pal <- colorNumeric(palette = "YlGnBu", domain = ga_sf$PCTCUMPVAX)

# Step 5: Create a bubble map with Leaflet
leaflet(ga_sf) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addCircles(
    ~st_coordinates(geometry)[, 1], ~st_coordinates(geometry)[, 2],  
    radius = ~PCTCUMPVAX * 500,  
    color = ~pal(PCTCUMPVAX),  
    weight = 1,
    opacity = 0.7,
    fillOpacity = 0.7,
    label = ~paste0(NAME, ": ", round(PCTCUMPVAX * 100, 2), "%"),  
    highlightOptions = highlightOptions(weight = 2, color = "#666", fillOpacity = 0.9, bringToFront = TRUE)
  ) %>%
  addLegend(pal = pal, values = ~PCTCUMPVAX, title = "Vaccination Rate %", position = "bottomright")
```

### Key Takeaways:
- Demonstrates a relationship between districts and vaccination rates.
- Highlights disparities in vaccine uptake.
- Guides policymakers to identify at-risk populations for outreach efforts.